---
title: "Northern gannets Rouzic 2022"
author: "Aurore Ponchon"
format: html
editor: visual
---

## 1- Importing data from Movebank

The dataset is stored on the online database Movebank. It can be downloaded directly from the web as long as the user has an account. Each time the dataset is downloaded, the last locations are added, as some of the devices are still emitting.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false
library(move)
library(here)
library(forcats)
library(tidyverse)
library(readxl)
library(ggplot2)
library(viridis)
library(gridExtra)
library(sf)
library(rgdal)
library(fields)
library(kableExtra)
library(conflicted)
library(adehabitatLT)


source(here("R","Functions.R"))

 # loginStored <- movebankLogin(username="AurorePonchon", password="AlbaNoir0*")
 # gannets<-searchMovebankStudies(x="gannet", login=loginStored)[1]
# metadata<-getMovebankStudy(study=gannets,login=loginStored)
# getMovebankSensorsAttributes(study=gannets,login=loginStored)
# reftable<-getMovebankReferenceTable(study=gannets,login=loginStored)
 # move.rouzic <- getMovebankData(study=gannets, login=loginStored,
 #                              removeDuplicatedTimestamps=TRUE)
 # save(move.rouzic,file=here("data","raw","gannets_rouzic_raw_2022.Rdata"))
# 
# leroySPDF <- as(move.rouzic,"SpatialPointsDataFrame")
# writeOGR(leroySPDF, dsn=here("data","NewlyCreatedData","shapefiles"), layer="all_raw_gannets", driver="ESRI Shapefile", overwrite_layer =T)

load(here("data","raw","gannets_rouzic_raw_2022.Rdata"))
```

## 2- Processing location data

Movebank datasets contain a lot of information that are not all necessary for the analysis so data were filtered. Some data from other files were added, such as individual serology. Then, we calculated distances and time between locations and added a few variables to summarize time spent on land and at-sea. We illustrate the raw data locations by some descriptive plots.

```{r}
#| echo: false
#| include: true
#| warning: false
#| message: false

colo_coord_rouzic<-data.frame(long=-3.436752, lat=48.899868)

#import of external individual data for serology
serol.rouzic<-read_excel(here("data","raw","Données Fous Rouzik 1O2022.xlsx"),
              sheet="Equipement GPS",
              n_max =16) %>% 
      dplyr::select(Balise,Résultats_test_Grippe_Aviaire) %>% 
       rename(id=Balise, sero=Résultats_test_Grippe_Aviaire) %>% 
       mutate(id=as.factor(id),
              sero=as.factor(sero)) %>% 
        mutate(sero = recode(sero,"Negatif"  = "Negative" , 
                                   "Positif" = "Positive"  ,
                                   "Inhibé" =  "Unknown" ,
                                   "Negatif au 1/10" = "LateNegative" ))
       
#select location data
locs.rouzic<- move.rouzic@data %>% 
    dplyr::select(tag_local_identifier,timestamp,location_long,location_lat,
           barometric_height, ground_speed) %>% 
    rename(id=tag_local_identifier,
           datetime=timestamp,
           long=location_long,
           lat=location_lat,
           alt=barometric_height,
           speed=ground_speed)  %>% 
    arrange(id,datetime) %>% 
    mutate(id=as.factor(id),site=as.factor("Rouzic")) %>% 
    left_join(serol.rouzic,by="id",keep=F) %>% 
    mutate(id=fct_drop(id))
    


ids<-unique(levels(locs.rouzic$id))
colo.ind<-tim.colors(length(ids))

#create new object with distance and time differences
rawLoc.rouzic<-NULL
#distance threshold set to 1km away from the colony to determine a trip
dist.threshold<-1    

#loop through each individual 
for (i in 1:length(ids)){
    ind.loc<-locs.rouzic %>% 
        dplyr::filter(id==ids[i]) %>% 
        mutate(distmax=0,
               distadj=0,
               totalpath=0,
               difftimemin=0,
               travelNb=0,
               onlandNb=0)
  
    print(c("Processing individual ", i))
      
  #loop through each line to calculate distance and time difference with previous location
for (l in 2:nrow(ind.loc)){
    
ind.loc[l,"distadj"]<-rdist.earth(ind.loc[l,c("long","lat")],
                                   ind.loc[l-1,c("long","lat")], miles = F)
ind.loc[l,"difftimemin"]<-difftime(ind.loc[l,"datetime"],
                                   ind.loc[l-1,"datetime"], units="mins")
ind.loc[l,"distmax"]<-rdist.earth(ind.loc[l,c("long","lat")],
                                   colo_coord_rouzic[c(1,2)], miles = F)
}

#define periods on land and at sea   
ind.loc<-define_trips(data=ind.loc,dist.min=dist.threshold)
rawLoc.rouzic<-rbind(rawLoc.rouzic,ind.loc)    
}

rawLoc.rouzic$trip.id<-paste(rawLoc.rouzic$id,rawLoc.rouzic$travelNb,sep=".")
save(rawLoc.rouzic,file=here("data","NewlyCreatedData","rawLocRouzic.Rdata"))


load(here("data","NewlyCreatedData","rawLocRouzic.Rdata"))

#Plot locations
distloc<-ggplot(rawLoc.rouzic,aes(x=datetime,y=distmax),group=id)+
    geom_point(aes(color=id),show.legend=F) +
     geom_line(aes(color=id),show.legend=F,size=0.8) +
#    scale_y_continuous(limits=c(0,400)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Maximal distance to colony (km)")

diffloc<-ggplot(rawLoc.rouzic,aes(x=datetime,y=difftimemin/60),group=id)+
    geom_point(aes(color=id),show.legend=F) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Time gaps (h)")

grid.arrange(distloc,diffloc,ncol=2)

#summarize all trips and periods on land by individuals
raw_trips1<-raw_trips_summary_ind(rawLoc.rouzic)
raw_land1<-raw_land_summary_ind(rawLoc.rouzic)

write.csv(raw_trips1,file=here("data","NewlyCreatedData","raw_trips_summary1.csv"),
          row.names = F,quote=F)
write.csv(raw_land1,file=here("data","NewlyCreatedData","raw_land_summary1.csv"),
          row.names = F,quote=F)

summy.trips<-raw_trips1 %>% 
    group_by(id) %>% 
    summarize(Nbtrip=n(),
              Mindurh=min(TripDurh),
              maxDiffh=max(maxDiffTimeh),
              minLocs=min(nlocs))

kbl(summy.trips,caption="Summary of raw trips") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


summy.land<-raw_land1 %>% 
    group_by(id) %>% 
    summarize(Nbland=n(),
              Mindurh=min(LandDurh),
              maxDiffh=max(maxDiffTimeh),
              minLocs=min(nlocs))

kbl(summy.land,caption="Summary of periods on land") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))


remove(move.rouzic,ind.loc,locs.rouzic,serol.rouzic)


##plot trip and land periods
dist<-ggplot(raw_trips1,aes(x=DateEnd,y=Distmaxkm),group=id)+
    geom_point(aes(color=id),show.legend=F) +
    scale_y_continuous(limits=c(0,200)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Maximal distance to colony (km)")

dur<-ggplot(raw_trips1,aes(x=DateEnd,y=TripDurh ),group=id)+
    geom_point(aes(color=id)) +
    scale_y_continuous(limits=c(0,10)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Trip duration (h")
grid.arrange(dist,dur,ncol=2)

correl<-ggplot(raw_trips1,aes(x=log(TripDurh),y=log(Distmaxkm)))+
    geom_point(aes(color=DateEnd)) +
    scale_colour_datetime(low=viridis(100)[1],high=viridis(100)[100]) +
    labs(x="log Trip duration (h)",y="log Maximal distance to colony (km)")
print(correl)

land<-ggplot(raw_land1,aes(x=DateEnd,y=LandDurh ),group=id)+
    geom_point(aes(color=id)) +
  #  scale_y_continuous(limits=c(0,10)) +
    scale_color_manual(values=colo.ind)+
    labs(x="Date",y="Periods on land (h)")
print(land)


##Parameters to define clean trips and filter locations
row.thres<- 4               #number of rows constituting a trip
dur.thres<-1         #minimum duration of a trip (in h)

clean.trip.nb<-raw_trips1 %>%
     dplyr::filter(TripDurh > dur.thres & nlocs >= row.thres) 

cleanLocs.rouzic<-rawLoc.rouzic %>%
    dplyr::filter(rawLoc.rouzic$trip.id %in% clean.trip.nb$trip.id)

save(cleanLocs.rouzic,file=here("data","NewlyCreatedData","clean_locs_rouzic.Rdata"))
```

## 3- Interpolation

As data has irregular time intervals between consecutive locations, we linearly interpolated data to have one location every 15min., as 96% of data have a time interval of less than 15min.We then compared each raw clean trips with interpolations.

```{r}

load(here("data","NewlyCreatedData","clean_locs_rouzic.Rdata"))


##Interpolation
reso<-60*15 #(in seconds)

loc.clean.interp<-interpol(data=cleanLocs.rouzic,time.int=reso)
loc.clean.interp$sero<-cleanLocs.rouzic$sero[match(loc.clean.interp$id,
                                                   cleanLocs.rouzic$id)]

save(loc.clean.interp,file=here("Data","NewlyCreatedData","clean_interp_rouzic.RData"))

##Ilustrate interoplations with raw data
#library(rnaturalearth)
#library(rnaturalearthdata)
library(maps)

for (a in 1:length(ids)){
    raw<-cleanLocs.rouzic %>% 
        dplyr::filter(id==ids[a] & lat > 47) 
    interp.ind<-loc.clean.interp %>% 
        dplyr::filter(id==ids[a]& lat > 47)
    

    plot(interp.ind$long,interp.ind$lat,pch=16,col="red",cex=1.2,main=ids[a])
    points(raw$long,raw$lat,pch=16,col="black",cex=0.8)
    points(colo_coord_rouzic$long,colo_coord_rouzic$lat,pch=17,col="blue",cex=1.5)
   legend("topleft","(x,y)",legend=c("raw","interp"),pch=16,col=c("black","red"),
          pt.cex=c(0.8,1.2))
}

```

## 4- Interpolation of at-sea locations

In order to determine an appropriate time interval, we investigate time difference between at-sea locations to identify potential gaps in the data and see when they occur. We also investigate the duration of at-sea trips.

```{r}
#| echo: false
#| include: true


```

After examination, 96% of data have a time interval of less than 15min so interpolation can be based on this time threshold.

```{r}
#| echo: false
#| include: true



```
