---
title: "4-Kernels"
author: "Aurore Ponchon"
format: html
editor: visual
---

# Calculating kernels

source(here("R","trip_functions.R"))

\# font_add_google(name="Rokkitt",family="rokkitt")

\# showtext_auto()

#bassrock for trips

load(here("Data","NewlyCreatedData","loc_clean_interp_bassrock.RData"))

load(here("Data","NewlyCreatedData","loc_clean_interp_rouzic.RData"))

```{r}

library(here)
library(tidyverse)
library(viridis)
library(gridExtra)
library(adehabitatHR)
library(sf)
library(rgdal)
library(ggspatial)
library(extrafont)

source(here("R","trip_functions.R"))
# font_add_google(name="Rokkitt",family="rokkitt")
# showtext_auto()

remotes::install_github("idmn/ggview")

#bassrock for trips
load(here("Data","NewlyCreatedData","loc_clean_interp_bassrock.RData"))
load(here("Data","NewlyCreatedData","loc_clean_interp_rouzic.RData"))

land<-readOGR(dsn=here("data","NewlyCreatedData","shapefiles"),layer="euro_afri_map")
proj4string(land)<-wgscrs
land_proj<-spTransform(land, projcrs)

mycol<-mako(4,begin=0.3,end=0.75)

coordinates(loc.interp.bassrock)<-c("long","lat")
proj4string(loc.interp.bassrock) <- wgscrs
bassrock_proj<-spTransform(loc.interp.bassrock, projcrs)

KUD.bassrock<- kernelUD(bassrock_proj[,"site"],same4all=T,h="href",grid=1000)
    KUDvol.bassrock <- getvolumeUD(KUD.bassrock)
    ver90.bassrock<-getverticeshr(KUDvol.bassrock ,90)
    ver50.bassrock<-getverticeshr(KUDvol.bassrock ,50)
  
coordinates(loc.interp.rouzic)<-c("long","lat")
proj4string(loc.interp.rouzic) <- wgscrs
rouzic_proj<-spTransform(loc.interp.rouzic, projcrs)

KUD.rouzic<- kernelUD(rouzic_proj[,"site"],h="href",same4all=T,grid=1000)
    KUDvol.rouzic <- getvolumeUD(KUD.rouzic)
    ver90.rouzic<-getverticeshr(KUDvol.rouzic ,90)
    ver50.rouzic<-getverticeshr(KUDvol.rouzic ,50)

 
    par(mfrow=c(1,2))    
plot(ver90.bassrock,col=mycol[2], border=mycol[2],main="Kernel Bass Rock")
plot(ver50.bassrock,col=mycol[1], border=mycol[1],add=T)
plot(land_proj,col="grey90",add=T)

plot(ver90.rouzic,col=mycol[4], border=mycol[4],main="Kernel Rouzic")
plot(ver50.rouzic,col=mycol[3], border=mycol[3],add=T)
plot(land_proj,col="grey90",add=T)

```

# Representing tracks over time for Bass Rock

```{r}

#limits for Bass Rock
# longmnbr=-5
# longmxbr=8
# latmnbr=53
# latmxbr=59.5

#calculate bottom left and top right projected coordinates
br<-data.frame(long=c(-5,9),lat=c(53,59.5)) %>%
    sf::st_as_sf(., coords = c("long","lat")) %>%
    sf::st_set_crs(wgscrs) %>%
    sf::st_transform(., projcrs)

longmnbr=365786
longmxbr=1176972
latmnbr=5874142
latmxbr=6647326

coord_bassrock<-colo_coord_bassrock %>% 
     sf::st_as_sf(., coords = c("long","lat")) %>% 
    sf::st_set_crs(wgscrs) %>% 
    sf::st_transform(., projcrs)

land_proj_bass<-land %>% 
    st_as_sf(.) %>% 
    # st_crop(., c(xmin = longmnbr, ymin = latmnbr, 
    #              xmax =longmxbr, ymax =latmxbr)) %>% 
   sf::st_transform(., projcrs)
 
maxdatebass<-loc.interp.bassrock %>% 
    mutate(day=format(datetime,format="%d/%m")) %>% 
    group_by(trip.id) %>% 
    summarize(Date=max(day))

sf_locs_bass <- loc.interp.bassrock  %>% 
    left_join(maxdatebass,by="trip.id") %>% 
    sf::st_as_sf(., coords = c("long","lat")) %>% 
    sf::st_set_crs(wgscrs) %>% 
    sf::st_transform(., projcrs) 

sf_lines_bass <- sf_locs_bass %>% 
  group_by(Date,trip.id) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING")  %>% 
  mutate(Date=as.Date(Date,format="%d/%m"))  %>% 
    arrange(Date)


bassgg<-ggplot()+
     geom_sf(data=land_proj_bass,fill="grey90",size=0.2) +   
     geom_sf(data = sf_lines_bass, mapping=aes(colour=Date),alpha=0.7) +
     geom_sf(data = coord_bassrock,shape=23, colour="black",fill="yellow", size=4) + 
     coord_sf(xlim = c(longmnbr,longmxbr), ylim=c(latmnbr, latmxbr),  expand = FALSE) +
     labs(tag="a)") +
     scale_colour_gradientn(colours=magma(length(unique(sf_lines_bass$Date))),
                            breaks = seq.Date(min(sf_lines_bass$Date),
                                       max(sf_lines_bass$Date),length.out=4),
                         labels = c("11/08", "31/08" ,"21/09" ,"12/10")) +
    annotation_scale(location = "br", width_hint = 0.3,
                     pad_x = unit(0.7, "cm"),bar_cols = "black") +
    annotation_north_arrow(location = "br", which_north = "true",
    style = ggspatial::north_arrow_nautical(), pad_y = unit(0.8, "cm")) +
    theme_bw() +
    theme(legend.position = "none",
          plot.margin = unit(c(0.5, 0.5, 0, 0), "lines"))

print(bassgg)

legbass<-g_legend(bassgg+theme(legend.position="left")+
    guides(color = guide_colorbar(reverse = TRUE)))

# write_sf(sf_lines_bass, dsn=here("data","NewlyCreatedData","shapefiles"),
#          driver="ESRI Shapefile",
#          layer="tracks_bassrock", append=F, delete_layer=T)
```

# Representing tracks over time for Rouzic

```{r}

#limits for Rouzic
# longmnr=-6
# longmxr=0
# latmnr=47.5
# latmxr=51

#calculate bttom left and top right projected coordinates
# br<-data.frame(long=c(-6,0),lat=c(47.5,51)) %>% 
#     sf::st_as_sf(., coords = c("long","lat")) %>% 
#     sf::st_set_crs(wgscrs) %>% 
#     sf::st_transform(., projcrs)

longmnr=274068.6
longmxr=710488.9
latmnr=5265093
latmxr=5654109

coord_rouzic<-colo_coord_rouzic %>% 
     sf::st_as_sf(., coords = c("long","lat")) %>% 
    sf::st_set_crs(wgscrs) %>% 
    sf::st_transform(., projcrs)


land_proj_rouz<-land %>% 
    st_as_sf(.) %>% 
    # st_crop(., c(xmin = longmnr, ymin = latmnr, 
    #              xmax =longmxr, ymax =latmxr)) %>% 
    sf::st_transform(., projcrs)   

maxdaterouz<-loc.interp.rouzic %>% 
    mutate(day=format(datetime,format="%d/%m")) %>% 
    group_by(trip.id) %>% 
    summarize(Date=max(day))

sf_locs_rouz<- loc.interp.rouzic  %>% 
    left_join(maxdaterouz,by="trip.id") %>% 
    sf::st_as_sf(., coords = c("long","lat")) %>% 
    sf::st_set_crs(wgscrs) %>% 
    sf::st_transform(., projcrs) 

sf_lines_rouz <- sf_locs_rouz %>% 
  group_by(Date,trip.id) %>%
  summarise(do_union = FALSE) %>%
  st_cast("LINESTRING") %>% 
  mutate(Date=as.Date(Date,format="%d/%m"))  %>% 
    arrange(Date)

rouzgg<-ggplot()+
     geom_sf(data = sf_lines_rouz, mapping=aes(colour=Date),alpha=0.7) +
     geom_sf(data=land_proj_rouz,fill="grey90",size=0.2) +   
     geom_sf(data = coord_rouzic,shape=23, colour="black",fill="yellow", size=4) + 
     coord_sf(xlim = c(longmnr,longmxr), ylim=c(latmnr, latmxr),  expand = FALSE) +
    labs(tag="b)") +
     scale_colour_gradientn(colours=viridis(length(unique(sf_lines_rouz$Date))),
                            breaks = seq.Date(min(sf_lines_rouz$Date),
                                       max(sf_lines_rouz$Date),length.out=4),
                         labels = c("24/08", "04/09" ,"16/09" ,"28/09")) +
      annotation_scale(location = "bl", width_hint = 0.3,
                     pad_x = unit(0.3, "cm"), bar_cols = "black") +
    annotation_north_arrow(location = "br", which_north = "true",
    style = ggspatial::north_arrow_nautical(), pad_y = unit(0.3, "cm")) +
    theme_bw() +
    theme(legend.position = "none",
          plot.margin = unit(c(0.5, 0.5, 0, 0), "lines"))

print(rouzgg)

legrouz<-g_legend(rouzgg+theme(legend.position="left")+
    guides(color = guide_colorbar(reverse = TRUE)))


# write_sf(sf_lines_rouz, dsn=here("data","NewlyCreatedData","shapefiles"),
#          driver="ESRI Shapefile",
#          layer="tracks_rouzic", append=F, delete_layer=T)


# tiff(here("outputs","Figure_maps.tif"),width=7000,height=4000,res=600,
#      compression="lzw")
# grid.arrange(bassgg,rouzgg,ncol=2,widths=c(1,1.1)) 
# 
#     dev.off()     
```

# Bearing of trips over time

```{r}
library(geosphere)
loc.rouz<-loc.interp.rouzic
loc.bass<-loc.interp.bassrock

#loc<-rbind(loc.rouz,loc.bass)

loc.rouz$Week<-as.factor(as.numeric(format(loc.rouz$datetime,
                                               format="%W"))-31)
#loc.rouz$Week<-factor(loc.rouz$Week, levels=c("8","7","6","5","4","3"))
loc.bass$Week<-as.factor(as.numeric(format(loc.bass$datetime,
                                               format="%W"))-31)
# loc.bass$Week<-factor(loc.bass$Week, 
#                       levels=c("10","9","8","7","6","5","4","3","2","1"))
loc.rouz$Bearing<-0
loc.bass$Bearing<-0
loc.rouz$Bearing<-(bearing(loc.rouz[,c("long","lat")],
                                         colo_coord_rouzic[,c("long","lat")],
                                         f=0))+180
loc.bass$Bearing<-(bearing(loc.bass[,c("long","lat")],                                                                   colo_coord_bassrock[,c("long","lat")],
                                         f=0))+180

mycol<-magma(10)

brk<-seq(0,360,30)
g1<-ggplot(loc.bass, aes(x = Bearing, group=Week)) +
  geom_histogram(aes(fill=Week), color = "gray10",breaks=brk) +
  scale_x_continuous(breaks = seq(0, 360, 30),limits=c(0,360)) +
     scale_y_continuous(breaks=seq(0,8000,2000),limits=c(0,8500))+
  scale_fill_manual(values = mycol) +
     labs(y="Nb of locations",x="",tag="c)")+
  coord_polar(theta="x",start = 0, direction = 1,clip="on") +
#  facet_grid(. ~ site) +
  theme_minimal() +
  theme( legend.position = "none",
        plot.margin = unit(c(0.5, 0, 0, 0), "lines"),
     axis.ticks.y = element_line(),
     axis.title.y = element_text(hjust=0.8))
  
  print(g1)  
  
 g2<-ggplot(loc.rouz, aes(x = Bearing,group=Week)) +
  geom_histogram(aes(fill=Week), color = "gray10",breaks=brk) +
  scale_x_continuous(breaks = seq(0, 360, 30),limits=c(0,360)) +
  scale_y_continuous(breaks=seq(0,5000,1000),limits=c(0,5000))+
  scale_fill_manual(values = viridis(6)) +
  coord_polar(theta="x",start = 0, direction = 1,clip="on") +
     labs(y="Nb of locations",x="",tag="d)")+
#  facet_grid(. ~ site) +
  theme_minimal() +
  theme( legend.position = "none",
          plot.margin = unit(c(0.5, 0, 0, 0), "lines"),
         axis.ticks.y = element_line(),
         axis.title.y = element_text(hjust=0.8))
  
  print(g2)  
  
  tiff(here("outputs","Figure_maps_bearing.tif"),width=6000,height=4000,res=600,
     compression="lzw")
   grid.arrange(bassgg,g1,legbass,
                            rouzgg ,g2 ,legrouz,
                                     nrow=2, ncol=3, widths=c(1,1,0.5))
  
  dev.off()
```
